---
title: "grade reflection"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Grade Reflection

# What Grade Did You Earn? I believe that I earned an A- in STA 518.

# How Did You Demonstrate That Grade? 
So for our grades, the course objectives is what we are to base our argument on. The reason I believe I earned an A- in this class would be based on the following:

Import, manage, and clean data: This one is pretty straight forward as i was able to import the results.csv. In the load phase I did get rid of all NA although I did not believe there were to be any, but this is one thing i like to make sure I do before any analysis is started. Also in my graph code chunks I use groupby, filter and summarise to manage data and extract the data I need. I do believe my understanding of this course objective is very good. This section came natural to me and I was able to learn rather quickly. I think I aced this part of the course objectives. I was not caught up by anything and learned the rules and parameters of functions very easily!
## Load Data

```{r echo = FALSE}

stats = read.csv("~/STA 518/Final Project/Portfolio/Final Proposal/Final Draft Portfolio/results.csv")

head(stats)

str(stats)

stats[is.na(stats)] = 0

summary(stats)

results = read.csv("~/STA 518/Final Project/Portfolio/Final Proposal/Final Draft Portfolio/results.csv")

head(results)

str(results)

summary(results)
```

Create graphical displays and numerical summaries of data for exploratory analysis and presentations.

Write R programs for simulations from probability models and randomization-based experiments: Welp, here we are.. the course objective that I was not able to complete. To be quite honest I have no idea how to do this or even where to start. This is the reason why I believe I did not earn a full A or A+ . I wish I was able to say on my reflection that I earned the full A, but I have no idea where to start with this or anything along those lines. Im hoping you understand that I just do not know how to do it. I could have tried, but like I said I dint even know where to start. 

Use source documentation and other resources to troubleshoot and extend R programs.

Write clear, efficient, and well-documented R programs.